### 2015-05-29 reports/paper/conclusion.tex
update thesis

\section{Conclusion}
In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

>>>>

\chapter{Conclusion}
In this paper, we proposed and built an inexpensive, portable yet reasonably accurate sound localization system with two microphone arrays. We analyzed different array architectures and demonstrated that a two array architecture achieves better accuracy compared to a single array system of similar size. By using cross-correlation output as the likelihood for different arrival time differences, each array generates a likelihood map for each possible location in the area. We demonstrated that by merging likelihood maps from the two arrays, we were able to achieve good localization accuracy (less than $3$ cm average error) in a local one meter by one meter region.

### 2015-05-29 reports/paper/prior.tex
update thesis

\chapter{Background}
Acoustic localization has been researched extensively in the literature. Localization techniques can be broadly categorized into Location Template Matching (LTM) and Time Difference of Arrival (TDOA) based approaches.

@@

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
\end{description}

>>>>

\chapter{Background}
Acoustic localization has been researched extensively in the literature. Localization techniques can be broadly categorized into Location Interaural Level Difference (ILD), Template Matching (LTM), and Time Difference of Arrival (TDOA) based approaches.

\section{ILD}
a

@@

Under the GCC framework, the arrival time difference $t_0$ between two signals $x_1(t)$ and $x_2(t)$ can be estimated as:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transforms of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to the cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}%[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divided by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. Part of magnitude information is included in delay estimation.
\end{description}

### 2015-05-29 reports/paper/system.tex
converted to thesis format
update thesis

\section{System Setup}
The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle. A micro-controller is attached to one of the vertices. Fig~\ref{fig:setup_array} shows a picture of the array setup. The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}. In this project, the micro-controller collects microphone data on all three channels for $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.

@@

To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$. The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the likelihood map $L(x,y)$ can be built by superposing the likelihood from each microphone pair:
\begin{eqnarray*}
L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Likelihood maps from two arrays can be combined into the final likelihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
, where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.

>>>>

\chapter{System Setup}
The end system has two arrays, each with three microphones mounted on the vertices of a $20$ cm equilateral triangle. A micro-controller is attached to one of the vertices. Fig~\ref{fig:setup_array} shows a picture of the array setup. The micro-controller used in this project is \emph{teensy 3.1}, which has $64$k RAM memory and the ADC is capable of sampling at $500$kHz~\cite{tdoa:micloc, sys:teensy}. In this project, the micro-controller collects microphone data on all three channels for $12$ milliseconds and then sends the recorded data to a computer through the USB port for localization.

@@

To handle the uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation output(equation~\ref{eq:gcc2}) as a measure of the likelihood of different arrival time differences. For each microphone array, we build a heatmap for the region. The intensity at each point represents the likelihood of it being the source. For each point on the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points on the grid and performing a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,$ and $m_3$, there are three microphone pairs: $m_1m_2,m_1m_3,$ and $m_2m_3$. The theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the likelihood map $L(x,y)$ can be built by superposing the likelihood from each microphone pair:
\begin{eqnarray*}
L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

Likelihood maps from two arrays can be combined into the final likelihood map:
\begin{equation}\label{eq:combine_l}
L(x,y) = L_1(x,y) L_2(x,y)
\end{equation}
where $L_1(x,y)$ and $L_2(x,y)$ represent the likelihood map from array $1$ and array $2$.
